---
title: "EPA Region 5 ambient air toxicity values"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
    highlight: tango
    theme: readable

---
  
<style type="text/css">
  body, td {font-size: 18px;}
  code.r{font-size: 18px;}
  pre {font-size: 18px} 
</style>
  

# Overview

## Table attributes

`Pollutant`    _The pollutant's common name. If more than one, use the first alphabetically._  
`CAS`          _The pollutant's unique ID assigned by the Chemical Abstract Service._  
`Mol_Wt`       _Molecular weight._  
`Acute_1hour_Benchmark`       _The pollutant's 1 hour concentration below which adverse health effects are not expected. Units of ug/m3._  
`Noncancer_Benchmark_Annual`       _The pollutant's annual concentration below which adverse health effects are not expected. Units of ug/m3._   
`Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000`       _The pollutant's annual concentration below which the additional cancer risk from a lifetime exposure is expected to be below 1 in 100,000. Units of ug/m3._   


# Load EPA and Region V state risk databases
```{r warning=F, message=F}

col_names <- c("Pollutant", "CAS", "Acute_1hour_Benchmark", "Noncancer_Benchmark_Annual", "Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000")

```

## EPA NATA risk values

NATA documentation available here - https://www.epa.gov/national-air-toxics-assessment/2011-nata-technical-support-document

```{r warning=F, message=F, results = 'hide'}
library(readxl)
library(tidyverse)

# Load
nata <- read_excel("..\\data\\Risk_values\\NATA_Pollutants_AppendixB_AppendixC.xlsx")
```

```{r warning=F, message=F}
# Add blank acute column
nata$Acute <- NA

# Select columns
names(nata)

nata <- nata[ , c(23,10,44,27,26)]

# Name columns
names(nata) <- col_names

# Clean CAS #'s
nata$CAS <- gsub("-", "", nata$CAS)

nata[grepl("Diesel", nata$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
nata$Noncancer_Benchmark_Annual <- 1000 * nata$Noncancer_Benchmark_Annual

nata$State <- "EPA"

```


## Wisconsin risk values

Documentation available here - http://dnr.wi.gov/topic/airquality/toxics.html

```{r warning=F, message=F, results = 'hide'}
# Load
df <- read_excel("..\\data\\Risk_values\\Wisconsin_CombinedNR445RevTables.xls", skip = 1)
```


```{r warning=F, message=F}
# Select columns
names(df)

df <- df[ , c(3,4,31,63,57)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)

#df[grepl("Diesel", df$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
#df$Noncancer_Benchmark_Annual <- 1000 * df$Noncancer_Benchmark_Annual
df$State <- "Wisconsin"

wisc <- df

```





## Michigan risk values

Documentation available here - http://www.deq.state.mi.us/itslirsl/

```{r warning=F, message=F, results = 'hide'}

# Load
df <- read_excel("..\\data\\Risk_values\\Mich-aqd-toxics-xcelitsl.xls")
```

```{r warning=F, message=F}
# Split acute and long-term noncancer columns
df <- df %>% mutate(Acute = ifelse(AQD_Averaging_Time == "1 hr", AQD_ITSL, NA))
df <- df %>% mutate(Acute = ifelse(AQD_Secondary_ITSL_Averaging_Time == "1 hr", AQD_Secondary_ITSL, Acute))

df <- df %>% mutate(Noncancer = ifelse(AQD_Averaging_Time == "annual", AQD_ITSL, NA))
df <- df %>% mutate(Noncancer = ifelse(AQD_Secondary_ITSL_Averaging_Time == "annual", AQD_Secondary_ITSL, Noncancer))


# Select columns
names(df)

df <- df[ , c(3,2,19,20,14)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)

#df[grepl("Diesel", df$Pollutant), ]$CAS <- "MDPM25"

df <- filter(df, CAS != 0)


# Convert cancer risk to 1 in 100,000 benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 10 * df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
# df$Noncancer_Benchmark_Annual <- 1000 * df$Noncancer_Benchmark_Annual

df$State <- "Michigan"

mich <- df

```


## Indiana risk values

Documentation available here - http://www.in.gov/idem/toxic/2343.htm

```{r warning=F, message=F, results = 'hide'}

# Load
df <- read_excel("..\\data\\Risk_values\\Indiana_Air_Toxics_Toxicity.xlsx")
```

```{r warning=F, message=F}
# Add blank acute column
df$Acute <- NA

# Select columns
names(df)

df <- df[ , c(1,2,7,5,3)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)

#df[grepl("Diesel", df$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
#df$Noncancer_Benchmark_Annual <- 1000 * df$Noncancer_Benchmark_Annual

df$State <- "Indiana"

ind <- df

```


## Combine risk tables
```{r warning=F, message=F}

# Combine
risks <- rbind(wisc[ ,1:6], mich[ ,1:6], nata[ ,1:6], ind[,1:6])

# Add units
risks$Units <- "ug/m3"

```






