---
title: "EPA Region 5 ambient air toxicity values"
output: 
  html_document:
    fig_width: 18
    fig_height: 9
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
    code_folding: hide
    highlight: tango
    theme: readable

---
  
<style type="text/css">
  body, td {font-size: 18px;}
  code.r{font-size: 18px;}
  pre {font-size: 18px} 
</style>
  

# Overview

## Table attributes

`Pollutant`    _The pollutant's common name. If more than one, use the first alphabetically._  
`CAS`          _The pollutant's unique ID assigned by the Chemical Abstract Service._  
`Mol_Wt`       _Molecular weight._  
`Acute_1hour_Benchmark`       _The pollutant's 1 hour concentration below which adverse health effects are not expected. Units of ug/m3._  
`Noncancer_Benchmark_Annual`       _The pollutant's annual concentration below which adverse health effects are not expected. Units of ug/m3._   
`Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000`       _The pollutant's annual concentration below which the additional cancer risk from a lifetime exposure is expected to be below 1 in 100,000. Units of ug/m3._   


# Load EPA and Region V state risk databases
```{r warning=F, message=F}
col_names <- c("Pollutant", "CAS", "Acute_1hour_Benchmark", "Noncancer_Benchmark_Annual", "Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000")
```

## EPA NATA risk values

NATA documentation available here - https://www.epa.gov/national-air-toxics-assessment/2011-nata-technical-support-document

```{r warning=F, message=F, results = 'hide'}
library(readxl)
library(tidyverse)
library(ggplot2)

# Load
nata <- read_excel("..\\data\\Risk_values\\NATA_Pollutants_AppendixB_AppendixC.xlsx")
```

```{r warning=F, message=F}
# Add blank acute column
nata$Acute <- NA

# Select columns
names(nata)

nata <- nata[ , c(23,10,44,27,26)]

# Name columns
names(nata) <- col_names

# Clean CAS #'s
nata$CAS <- gsub("-", "", nata$CAS)

nata[grepl("Diesel", nata$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / nata$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
nata$Noncancer_Benchmark_Annual <- 1000 * nata$Noncancer_Benchmark_Annual

nata$State <- "EPA"

```


## Wisconsin risk values

Documentation available here - http://dnr.wi.gov/topic/airquality/toxics.html

```{r warning=F, message=F, results = 'hide'}
# Load
df <- read_excel("..\\data\\Risk_values\\Wisconsin_CombinedNR445RevTables.xls", skip = 1)
```


```{r warning=F, message=F}
# Select columns
names(df)

df <- df[ , c(3,4,31,63,57)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)

#df[grepl("Diesel", df$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

# Convert non-cancer reference concentrations (RFCs) to ug/m3
#df$Noncancer_Benchmark_Annual <- 1000 * df$Noncancer_Benchmark_Annual
df$State <- "Wisconsin"

wisc <- df

```


## Michigan risk values

Documentation available here - http://www.deq.state.mi.us/itslirsl/

```{r warning=F, message=F, results = 'hide'}

# Load
df <- read_excel("..\\data\\Risk_values\\Mich-aqd-toxics-xcelitsl.xls")
```

```{r warning=F, message=F}
# Filter missing
df <- filter(df, !is.na(CAS_Number), CAS_Number != 0)

# Split acute and long-term noncancer columns
df <- df %>% group_by(CAS_Number) %>% mutate(Acute = (AQD_Averaging_Time == "1 hr") * AQD_ITSL)

df <- df %>% group_by(CAS_Number) %>% mutate(Acute2 = (AQD_Secondary_ITSL_Averaging_Time == "1 hr") * AQD_Secondary_ITSL)

df <- df %>% group_by(CAS_Number) %>% mutate(Acute = max(c(Acute, Acute2, 0), na.rm=T))

df <- df %>% group_by(CAS_Number) %>% mutate(Noncancer = (AQD_Averaging_Time == "annual") * AQD_ITSL)

df <- df %>% group_by(CAS_Number) %>% mutate(Noncancer2 = (AQD_Secondary_ITSL_Averaging_Time == "annual") * AQD_Secondary_ITSL)

df <- df %>% group_by(CAS_Number) %>% mutate(Noncancer = max(c(Noncancer, Noncancer2, 0), na.rm=T))


# Select columns
names(df)

df <- df[ , c(3,2,19,21,14)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)
df[grepl("diesel", df$Pollutant), ]$CAS <- "MDPM25"


# Convert cancer risk to 1 in 100,000 benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 10 * df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)

df$State <- "Michigan"

mich <- df

```


## Indiana risk values

Documentation available here - http://www.in.gov/idem/toxic/2343.htm

```{r warning=F, message=F, results = 'hide'}

# Load
df <- read_excel("..\\data\\Risk_values\\Indiana_Air_Toxics_Toxicity.xlsx")
```

```{r warning=F, message=F}
# Add blank acute column
df$Acute <- NA

# Select columns
names(df)

df <- df[ , c(1,2,7,5,3)]

# Name columns
names(df) <- col_names

# Clean CAS #'s
df$CAS <- gsub("-", "", df$CAS)
df[grepl("Diesel", df$Pollutant), ]$CAS <- "MDPM25"

# Convert cancer risk to benchmark concentration
df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 <- ifelse(df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000 > 0, 1E-05 / df$Cancer_Benchmark_Addl_Lifetime_Cancer_Risk_of_1in100000, NA)


df$State <- "Indiana"

ind <- data.frame(df, stringsAsFactors = FALSE)

```



## Load Minnesota risk values

Documentation available here - https://www.pca.state.mn.us/air/air-emissions-risk-analysis-aera-rass-and-qchi-spreadsheets  


```{r warning=F, message=F}
minn <- read.csv("..\\data\\air_benchmarks.csv", check.names = F, stringsAsFactors = F)

minn <- minn[ , c(2,1,3,5,6)]

# Save column names
minn_names <- names(minn)

# Update column names to join with other states
names(minn) <- col_names

# Clean CAS #'s
minn[grepl("Diesel", minn$Pollutant), ]$CAS <- "MDPM25"

minn$CAS2 <- minn$CAS

minn$CAS <- gsub("-", "", minn$CAS)

# Assign state
minn$State <- "Minnesota"

```

# Combine risk tables

```{r warning=F, message=F}

# Combine
risks <- rbind(wisc[ ,1:6], mich[ ,1:6], nata[ ,1:6], ind[,1:6])

# Missing pollutants
risks_miss <- filter(risks, !CAS %in% minn$CAS)

# Filter to Minnesota pollutants
risks <- filter(risks, CAS %in% minn$CAS)

# Add Minnesota
#minn <- filter(minn, CAS %in% risks$CAS)
risks <- rbind(risks, minn[ , c(1:5, 7)])

risks <- left_join(risks, minn[ , c("CAS", "CAS2")]) 

risks$CAS  <- risks$CAS2

risks$CAS2 <- NULL

# Count number of states with risk values
risks <- risks %>% group_by(CAS) %>% mutate(count = n())

# Updata column names to align with web application
names(risks)[1:5] <- minn_names

# Add units
#risks$Units <- "ug/m3"

# SAVE table
write.csv(risks[ , 1:6], "..//data//air_benchmarks_region5.csv")

```


# Plot comparisons
```{r warning=F, message=F}

compare_pol <- c("50-00-0", "106-99-0", "7440-02-0", "MDPM25", "107-02-8")

# Remove pollutants measured by only one state
risks <- filter(risks, count > 2)  

# Plot risk values by state
for(i in compare_pol) {
  df <- filter(risks, CAS == i)
  
  p <- ggplot(df, aes(State, `Chronic Non-cancer Reference Conc (ug/m3)`)) + geom_bar(stat="identity") + labs(title= paste(df$Pollutant[1], "Noncancer threshold"), subtitle=i)
  
  print(p)
  
   p <- ggplot(df, aes(State, `Chronic cancer risk of 1E-5 Air Conc (ug/m3)`)) + geom_bar(stat="identity") + labs(title= paste(df$Pollutant[1], "Cancer benchmark"), subtitle=i)
  
  print(p)

}
                          
```



